Carrier phase and amplitude estimation for phase shift keying using pilots and data (Appendix)

### 2012-08-16 paper1.tex
modified the 2 parters.  The second paper is not just appendix.  I think this makes alot of sense.

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data 1: \\ Estimators and simulations}
\author{Robby McKilliam, Andre Pollok, Bill Cowley  
\thanks{
Supported under the Australian Government’s Australian Space Research Program.
Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
}}

@@

The paper is organised as follows BLERG.

@@

\section{The Gaussian noise case}

@@

\normalsize
\appendix

>>>>

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
\author{Robby McKilliam, Andre Pollok, Bill Cowley  
\thanks{
Supported under the Australian Government’s Australian Space Research Program.
Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
}}

@@

The paper is organised in the following way.  Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the coherent case, when pilot symbols and data symbols exist.  Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.  Section~\ref{sec:stat-prop-least} states two theorms that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.  We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.  However, $\hat{\rho_0}$ is not a consistent estimator of the amplitude $\rho_0$.  The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.  Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} provide proofs of the theorems stated in Section~\ref{sec:stat-prop-least}.  Due to space constraints, the lemmas needed by these proofs have been placed in the supplementary paper~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.  In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.  In this case, our expressions for the asymptotic variance can be simplified.  Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.  These simulations agree with the derived asymptotic properties.

@@

\section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}

Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
 \begin{align*}
SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
\end{align*}
where
\[
Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
\]
and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.  Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$, 
\begin{equation}\label{eq:hatrhoZ}
\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
\end{equation}
where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
\[
SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
\]

We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.  Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
\begin{align*}
y_i &= a_0 s_i + w_i \\
&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
&= a_0 s_i R_i e^{j \Phi_i} \\
&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
\end{align*}
%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
\begin{align*}
\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
&= \round{ \lambda + \Phi_i + \angle{s_i} },
\end{align*}
where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
\begin{align}
 y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
\end{align}
since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
\[
y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
\]
Now,
\[
Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
\]
and
\[
\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
\] 
Let 
\begin{equation}\label{eq:GLdefn}
G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
\end{equation}
and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.  Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.  We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.  The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.

Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
\[
\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
\]
and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
\[
\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
\]
As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.  Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.  Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.  Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.  Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$. So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.  Part 1 of Theorem~\ref{thm:consistency} follows.

It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
\begin{equation}\label{eq:rhoGLZ}
\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
\end{equation}  
Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.

\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}

We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.  Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.  Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.  The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.  This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$. Define the function
\begin{equation}\label{eq:RLdef}
R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\end{equation}
Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.  Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
\begin{align*}
0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
\end{align*}
Lemma~\ref{lem:Qconv} shows that
\[
\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.  Lemma~\ref{lem:empiricprocc} shows that
\[
\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
\]
It follows from the three equations above that,
\[
0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
\]
and rearranging gives,
\[
\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
\]
Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.  It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
\[
\frac{pA_1 + dA_2}{(p + Hd)^2}.
\]

We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
\begin{align*}
\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
\end{align*}
Lemma~\ref{lem:empiricprocforrho} shows that 
\[
\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
\]
where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
\[
%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
\]
It follows that
\[
%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
\]
Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.  Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.  Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
\[
\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
\]
as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.

\section{The Gaussian noise case}\label{sec:gaussian-noise-case}

@@

### 2012-08-16 paper2.tex
modified the 2 parters.  The second paper is not just appendix.  I think this makes alot of sense.

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data 2: Proofs}
\author{Robby McKilliam, Andre Pollok, Bill Cowley 
\thanks{
Supported under the Australian Government’s Australian Space Research Program.
Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
}}

@@

\section{Introduction}

\section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}

Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
 \begin{align*}
SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
\end{align*}
where
\[
Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
\]
and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.  Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$, 
\begin{equation}\label{eq:hatrhoZ}
\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
\end{equation}
where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
\[
SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
\]

We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.  Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
\begin{align*}
y_i &= a_0 s_i + w_i \\
&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
&= a_0 s_i R_i e^{j \Phi_i} \\
&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
\end{align*}
%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
\begin{align*}
\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
&= \round{ \lambda + \Phi_i + \angle{s_i} },
\end{align*}
where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
\begin{align}
 y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
\end{align}
since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
\[
y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
\]
Now,
\[
Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
\]
and
\[
\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
\] 
Let 
\begin{equation}\label{eq:GLdefn}
G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
\end{equation}
and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.  Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.  We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.  The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.

Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
\[
\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
\]
and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
\[
\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
\]
As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.  Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.  Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.  Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.  Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$. So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.  Part 1 of Theorem~\ref{thm:consistency} follows.

It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
\begin{equation}\label{eq:rhoGLZ}
\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
\end{equation}  
Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.

@@

\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}

We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.  Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.  Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.  The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.  This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$. Define the function
\begin{equation}\label{eq:RLdef}
R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\end{equation}
Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.  Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
\begin{align*}
0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
\end{align*}
Lemma~\ref{lem:Qconv} shows that
\[
\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.  Lemma~\ref{lem:empiricprocc} shows that
\[
\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
\]
It follows from the three equations above that,
\[
0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
\]
and rearranging gives,
\[
\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
\]
Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.  It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
\[
\frac{pA_1 + dA_2}{(p + Hd)^2}.
\]

We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
\begin{align*}
\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
\end{align*}
Lemma~\ref{lem:empiricprocforrho} shows that 
\[
\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
\]
where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
\[
%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
\]
It follows that
\[
%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
\]
Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.  Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.  Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
\[
\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
\]
as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.

@@

\section{Conclusion}

\normalsize
\appendix

@@

\subsection{An empirical process result}

>>>>

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data (Appendix)}
\author{Robby McKilliam, Andre Pollok, Bill Cowley 
\thanks{
Supported under the Australian Government’s Australian Space Research Program.
Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
}}

@@

\section{Introduction to this appendix}

Here we compile lemmas required for the proof almost sure convergence (Theorem~\ref{thm:consistency}) and asymptotic normality (Theorem~\ref{thm:normality}) from the paper~\cite{McKilliam_leastsqPSKpilotsdata_2012}.  This document is not self contained and it is intended that the reader start at~\cite{McKilliam_leastsqPSKpilotsdata_2012}.  The equation reference numbers in this document start at (?).  All reference numbers less than (?) refer back to~\cite{McKilliam_leastsqPSKpilotsdata_2012}.  This document is separated into 3 sections.  Section~\ref{sec:proof-almost-sureappendix} contains those lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}).  Section~\ref{sec:proof-asympt-normappendix} contains those lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}).  Section~\ref{sec:an-empirical-process} contains proofs for lemmas that arise during Section~\ref{sec:proof-asympt-normappendix}.  These involve the proof of~\emph{tightness} for certain empirical processes~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.

\section{Lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sureappendix}

@@

\section{Lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-normappendix}

@@

@@

\section{An empirical process result}\label{sec:an-empirical-process}
